Trending assessment for nowcasts, measurements, and forecasts

### 2024-02-20 sections/03-trending.tex
Update on Overleaf.

Confidence intervals can be issued to account for the estimation uncertainty of the measures above.
Bootstrapping confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
Whereas classical confidence intervals are computed from parametric assumptions on the underlying data, new samples are drawn with replacement from the dataset in bootstrapping.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
The basic approach balances those two aims.
We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time.
We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
The considered datasets are again outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
\todo[inline]{Add evaluation when computation is finished}

>>>>

Confidence intervals can be issued to account for the estimation uncertainty of the measures above.
Bootstrapping confidence intervals are a nonparametric technique based on resampling~\parencite[for introductions see][]{Hesterberg2011,Bittmann2021}.
Whereas classical confidence intervals are computed from parametric assumptions on the underlying data, new samples are drawn with replacement from the dataset in bootstrapping.
The confidence interval is then computed based on the bootstrap samples.
We examine three methods for bootstrapping here: the intuitive percentile and the more sophisticated basic and \ac{bca} method.
In the \textit{percentile} approach, the confidence interval for level $\alpha$ is built directly from the bootstrap estimators' empirical distribution.
The \textit{basic} approach computes the confidence interval based on the non-bootstrapped estimate using the bootstrapped quantile deviations~\parencite{Davison1997}.
The \ac{bca} method modifies the quantiles of the empirical bootstrap distribution by a bias and an acceleration parameter~\parencite{Efron1987}.
Typically, the percentile approach needs larger datasets and provides an easy and fast estimate, while the \ac{bca} is computationally expensive but needs smaller datasets for reasonable confidence intervals.
The basic approach balances those two aims.
We compare the approaches in a small synthetic data study concerning their small-dataset behavior and computation time.
We vary the number of available time steps $T$ to be a typical time-series value, such as 30 for daily data in a month, 52 for weekly data, 168, 365, 720, and 1024.
The considered datasets are again outlined in Appendix~\ref{sec:app-trending-data-generation}, the first dataset with asymmetric dependence.
In the calculations, the \verb|scipy| package's implementation of bootstrap confidence intervals is used~\parencite{Virtanen2020}.
The prescribed confidence level is 90 \%, and the number of bootstrap samples is $10,000$.
The share of confidence intervals covering the true values per method and $T$ are shown in Table~\ref{tab:trending_bootstrap}.
The true values of the accuracy are computed based on a dataset of size $10^8$, yielding 0.7501 and 0.7700 for the two datasets.
The computation times per method and dataset are shown in Figure~\ref{fig:trending_bootstrap_time}.
For the small sample sizes up to $T = 168$, only the \ac{bca} method keeps the confidence interval size and yields slightly wider confidence intervals. 
The method's results do not differ for the larger sample sizes. 
The computation time for the \ac{bca} method is slightly larger than for the other methods, but all methods have a moderate computation time.
